Omics-based Single Step Trait Prediction

### 2017-03-28 paper/manuscript.tex
Add minor points to discussion

\section*{Discussion}
\Genetics2level{Advantages of marker effect models}
% Describe the many hacks that have been developed to tackle problems with BLUP
% models.

@@

\Genetics2level{Superiority of single step models}

\Genetics2level{Coverage of the genetic space}
Animal populations are so large that it does not truly matter, which animals are 
being genotyped, but in plant breeding the situation is different and we would 
be interested in determining which individuals should be complemented with 
information on another predictor.

Prediction of phenotypic performance is particularly crucial in hybrid breeding
programs because the number of potential selection candidates is $n^{2}$ where
$n$ corresponds to the number of parent lines.

Given recent results on the utility of downstream 'omics' predictors in a study
on hybrid prediction in a European hybrid maize panel \cite{Westhues2017}, our 
objectives were i) to explore the benefits of a single step procedure for the 
same material and ii) to examine the influence on the composition of the set of
genotypes that is covered by more than one predictor on the success of single
step prediction.

% Special situation in hybrid breeding.
Hybrid breeding --- with noteworthy commercial applications in pig, chicken and 
maize breeding --- is a particularly challenging application field for
prediction tools.
Here, $2n$ parent individuals from two genetically distinct heterotic groups are
crossed to each other, yielding $n^{2}$ potential hybrid progeny that would
require intensive field testing.
In medium-sized plant breeding programs, the advent of the doubled haploid (DH) 
technology \cite{Wedzony2009} now allows for an annual production of 1,000 
parent lines in each heterotic group, amounting to $10^{6}$ putative hybrid
progeny.

So far, its applications in hybrid populations are limited to pig breeding
\cite{Xiang2015,Xiang2016,Tusell2016}.
% Are there applications in chicken breeding? (Hints in Christensen et al.
% 2014, GSE).

>>>>

\section*{Discussion}

\Genetics2level{Advantages of marker effect models}
% Describe the many hacks that have been developed to tackle problems with BLUP
% models.

@@

\Genetics2level{Superiority of single step models}
\subsection{Inbred lines}
Predictive abilities for the maize diversity panel are not in complete
agreement with those reported by \citeNP{Guo2016}.
Reasons for this may be threefold: 
1) In our study, we reduced the dataset to the 211 tropical/subtropical lines 
and excluded all other genotypes whereas \citeNP{Guo2016} used genotypes from 
four subpopulations and only a subset of 157 genotypes from the set of 
tropical/subtropical lines while correcting for population structure by
including a fixed effect in the model for predicting agronomic performance.
2) We applied quality checks for the predictor data after generating the subset
of 211 inbred lines, whereas \citeNP{Guo2016} applied the quality checks to 368
genotypes.
3) While \citeNP{Guo2016} used five-fold cross-validation with 500 repetitions,
we employed a leave-one-out cross-validation scheme (LOOCV).

\subsection{Maize hybrids}
Hybrid breeding --- with noteworthy commercial applications in pig, chicken and 
maize breeding --- is a particularly challenging application field for
prediction tools.
Here, $2n$ parent individuals from two genetically distinct heterotic groups are
crossed to each other, yielding $n^{2}$ potential hybrid progeny that would
require intensive field testing.
In medium-sized plant breeding programs, the advent of the doubled haploid (DH) 
technology \cite{Wedzony2009} now allows for an annual production of 1,000 
parent lines in each heterotic group, amounting to $10^{6}$ putative hybrid
progeny.
So far, its applications in hybrid populations are limited to pig breeding
\cite{Xiang2015,Xiang2016,Tusell2016}.
% Are there applications in chicken breeding? (Hints in Christensen et al.
% 2014, GSE).

\begin{itemize}
  \item Generally, it is important that, on the same set of genotypes, the 
    incomplete predictor yields larger predictive abilities than the complete 
    predictor.
  \item In hybrid breeding, the probability of successfully selecting the best
    observed genotypes based on the best predicted candidates is a strongly 
    convex function of the predictive ability.
    Hence, even the minor gains when using transcriptomic and genomic data in a
    single step might justify additional investments in RNA-seq at least of a
    subste of genotypes that represents the genetic space of the breeding 
    material well.
  \item For the hybrid data, the predictive abilities of genomic information
    for the whole set of available parent lines rose considerably for several
    agronomic traits.
    Hence, the inclusion of trancsriptomic data could not further improve upon
    this result.
\end{itemize}

\Genetics2level{Coverage of the genetic space}
Animal populations are so large that it does not truly matter, which animals are 
being genotyped, but in plant breeding the situation is different and we would 
be interested in determining which individuals should be complemented with 
information on another predictor.
If one can cover the genetic space sufficently with the incomplete predictor,
